---
title: "Synthetic Data Simulation Results"
output: pdf_document
date: "2023-12-04"
---

```{r setup, include=FALSE}
library(pascsim)
library(tmle3sim)
library(ggplot2)
library(data.table)
library(knitr)
library(stringr)
library(lubridate)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


## Simulation Summary

LTMLE generated quite a few errors, reducing results given available time. Below 
are counts of completed runs and runs with errors for each simulation setting.

```{r sim_iter}

# results <- load_results("Results_backup/")

results <- load_results()
results <- rbindlist(results, fill = TRUE)

if(!("child_simulation_uuid"%in%names(results))){
  results$child_simulation_uuid <- NA
}

# results <- results[is.na(child_simulation_uuid)]
valid_n <- c(1e3,1e4,1e5)
results <- results[!is.na(n)&(n%in%valid_n)]
sim_names <- c("SyntheticData", "PascSim")
data_types <- c("Synthetic", "MC")
results[,data := data_types[match(simulation_name, sim_names)]]

est_names <- c("pascSummary","SyntheticDGP")
est_labels <- c("Marginal","DGP")
results[,estimator := est_labels[match(estimator_name, est_names)]]

# find errors by looking for missing results by seed
results[,error:=regime=="error"]
errors <- results[,list(data = data[1],
                        estimator = setdiff(est_labels, unique(estimator)),
                        n = n[1]), 
                  by = list(seed)]
errors <- errors[!is.na(estimator),]
errors[,error:= TRUE]
results_w_errors <- rbind(results, errors, fill = TRUE)
iter_counts <- results_w_errors[,list(count = length(unique(seed))),by=list(data, estimator, learners, n, error)]
iter_counts[,error:=ifelse(error, "error", "completed")]
wide <- dcast(iter_counts, data + estimator + learners ~ error + n, fill = 0, value.var = "count")
wide <- wide[!(data=="Synthetic"&estimator=="DGP")]
kable(wide)
```
ignore the error counts here (they're for sims we don't expect)

```{r sim_time}

iter_times <- results[,list(time = mean(runtime)),by=list(learners, data, estimator, n)]
format_duration <- function(x){
  dur <- round(as.period(dseconds(x)))
  
  sub_day <- sprintf("%02d:%02d:%02d", hour(dur), minute(dur), second(dur))
  formatted <- ifelse(day(dur), sprintf("%d day %s", day(dur), sub_day), sub_day)
}
x <- iter_times$time
iter_times[,duration:=format_duration(time)]
wide <- dcast(iter_times, data + estimator + learners ~ n, value.var = "duration")
kable(wide)
```

```{r calc_perf}
load("psi_0.Rdata")
results[is.na(effect_size), effect_size:=0.1]
point_estimates <- c("covid", "pasc", "death", "vax",
                     "metformin", "paxlovid")

# recalculate SE for marginal estimates
se_p <- function(x,n){
   x <- pmin(pmax(x,1/n),1-(1/n))
   sqrt(x*(1-x))/sqrt(n)
}
results[period%in%point_estimates,se:=se_p(mean,n)]
results[,regime:=as.character(regime)]

results <- merge( results, psi_0, by = c("period","regime","effect_size"))
# 
# 
# # look at performance relative to parent sample ests
# parent_seeds <- unique(results$parent_seed)
# parent_results <- results[seed%in%parent_seeds,
#                           list(seed = seed, period = period, n = n,
#                                regime = regime, effect_size = effect_size,
#                                psi_0 = mean, `se(psi_0)`=se)]
# child_results <- results[!is.na(parent_seed),
#                          list(seed = parent_seed, period = period,
#                               regime = regime,
#                               mean = mean, se = se)]
# child_results$data <- "Synth Relative"
# combined <- merge(parent_results, child_results, by=c("seed","period","regime"))

# results <- rbind(results, combined, fill = TRUE)

# truncate giant SEs and throw out missing values
# TODO: figure out why these are happening (Positivity)
results <- results[!is.na(mean)&!is.na(se)]

results[, se:=ifelse(se>quantile(se,0.9), quantile(se,0.9), se), by=list(regime, period)]

results[,bias:=mean-psi_0]
results[,var:=(mean-mean(mean))^2, by=list(data, period, regime, n)]
results[,mse:=(mean-psi_0)^2]
results[,lower:=mean-1.96*se]
results[,upper:=mean+1.96*se]
results[,coverage:=data.table::between(psi_0, lower,  upper)]
results[,ci_length:=upper-lower]
results[,power:=!data.table::between(0, lower,  upper)]


conditions <- c("data","period","regime","n", "learners")
metrics <- c("bias", "var", "mse", "coverage", "ci_length", "power")
metric_labels <- c("Bias", "Variance", "MSE", "Coverage", "CI Length", "Power")
metric_labels <- factor(metric_labels, levels = metric_labels)
perf <- performance_summary(results, conditions, metrics)
perf[,label:=metric_labels[match(metric, metrics)]]
# display absolute bias instead of bias w/ sign for easier comparison
suppressWarnings(perf[,lower2:=lower])
perf[metric=="bias", lower:=ifelse(value<0,-1*upper, lower)]
perf[metric=="bias", upper:=ifelse(value<0,-1*lower2, upper)]
perf[metric=="bias", value:=abs(value)]
perf$lower2 <- NULL

```

\newpage
## Point Estimate Performance

Point estimate performance was degraded, improving somewhat at 
larger sample sizes. This suggests a need for estimation with more data 
adaptive models.


```{r perf_pe, results = "asis"}
perf_pe <- perf[period%in%point_estimates&metric!="power"]
perf_pe[,regime:=as.numeric(regime)]
perf_pe <- perf_pe[metric!="power"]
sample_sizes <- sort( unique(perf_pe$n))
for(n_i in sample_sizes){
  cat(sprintf("  \n### n = %d  \n", n_i))
  p <- ggplot(perf_pe[n==n_i], aes(x=regime, y= value,
                                   ymin = lower, ymax = upper, 
                                   color = factor(data):factor(learners), fill = factor(data):factor(learners)))+
              geom_line()+geom_ribbon(alpha=0.2)+
              facet_grid(label~period, scales="free_y") + theme_bw() +
              xlab ("Time (days)") + ylab("Metric")
  print(p)
  cat("  \n")
}
```
# Bias
```{r bias, results = "asis"}
bias <- perf_pe[label=="Bias"]
for(n_i in sample_sizes){
  cat(sprintf("  \n### n = %d  \n", n_i))
  p <- ggplot(bias[n==n_i], aes(x=regime, y= value,
                                   ymin = lower, ymax = upper, 
                                   color = factor(learners), fill = factor(learners)))+
              geom_line()+geom_ribbon(alpha=0.2)+
              facet_grid(data~period) + theme_bw() +
              xlab ("Time (days)") + ylab("Metric")
  print(p)
  cat("  \n")
}
```
